Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-03-27 manuscript.Rmd
Clean through to methods

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

@@

# SUBMITTED DATA IS BASED ON THE FOLLOWING IDS
#9908  9909  9910  9911 10201 10203 10204 10301 10501 10601 10701 10702 10703 10704 10802 10901 11001 11002 11004 11102 11103 11104 11202 11501 11601 11901 12001 12002 12101 12302 12303 12501 12502 13001

@@

In this paper, we present an inertial sensing method to classify infant body position from moment to moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders. We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022], the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings. Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion [@AiraksinenRasanen2020] and infant-caregiver contact [@YaoPlotz2019]. 
<!-- Accompanying our paper, we have created a database [CITE] containing synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.  -->

@@

The LENA&reg; recorder is a commercial device that is worn by infants in a custom shirt pocket; the recorder has sufficient battery life and storage to record for an entire day. Closed-source LENA&reg; algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the infant. Other long-form audio methods rely on custom-built recorders [@WassPhillips2022], apply alternative classification algorithms to LENA&reg; data [@RasanenSeshadri2020; @MichelettiYao2022], or manually transcribe audio recorded by LENA&reg; devices to improve accuracy or identify behaviors beyond the built-in categories [@Bergelson2019Corpus; @MendozaFausey2021].

Long-form audio recordings have had a transformational impact on language development research by allowing researchers to characterize opportunities for learning in daily life. For example, measuring the amount of speech heard by infants in the home [@Weisleder2013] or in a daycare setting [@PerryPrince2018] revealed how individual differences in speech input predict later vocabulary. Full-day language recording synchronized with other data sources allows researchers to identify how auditory input and vocal production interact with other processes. Beyond individual differences in aggregated data, long-form recordings can be used to determine the temporal schedule of experiences. For example, infants' daily experiences hearing music are clustered in time, with "bursty" episodes of music presence separated by relatively long periods during which music is absent [@MendozaFausey2022]. In addition, synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences. Linking LENA&reg; speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during periods that they were restrained in devices such as swings, exersaucers, and car seats [@MalachowskiSalo2023]. A custom-built wearable ECG and audio recorder allowed @WassPhillips2022 to discover that infant arousal increases the likelihood of infant vocalization across the day.

We identified five key features of long-form audio methods that should be replicated in analogous studies of motor behavior. First, wearable audio recorders are *mobile*. Measurement is not limited to a particular room because the recording device travels with the participant. Data are recorded to onboard device memory, so participants do not need to be in range of a receiver. Second, wearable audio recording is *unobtrusive*. Participants' reactivity to observation, such as from a video camera, may influence behavior. For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA&reg; device [@Bergelson2019Input]. Third and fourth, recordings capture *real-time data* over a *full day*. The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants. Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual [e.g., @WassPhillips2022]. Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) [@KadookaCaufield2021; @Tamis-LeMondaCustode2018]. "Burstiness" of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent [@BarbaroFausey2022; @WarlaumontSobowale2021]. Fifth, *automatic classification* means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription. But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased. An independent assessment of the LENA&reg; algorithms found mixed results about the accuracy of different outcomes. For example, correlations between human transcribed counts of adult words and child vocalizations against LENA&reg;'s automatic counts were strong, *r* = .698 and *r* = .649, respectively [@CristiaLavechin2020]. For other metrics, such as the number of "conversational turns" between the child and communicative partners and the identification of male speakers, agreement was poor. 
<!-- EVENTUALLY CITE NEW EMA HERE -->

Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day. In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior.

@@

Video and survey methods are the current state-of-the-art in assessing motor behavior. Although each method has complementary advantages and disadvantages for characterizing infants' everyday motor experiences, neither method on its own can provide comparable data to the long form audio recordings reviewed in the previous section.

@@

Accordingly, the overarching goal of the current study is to test the validity of long-form body position recording in the home during unsupervised, everyday behavior. Supervised recordings from past work [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022], whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors. Prior work focused on "free play", in which caregivers were asked to play with the infant without restraining the infant or otherwise shaping how they could move. However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position. Indeed different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to be counted as equal. Moreover, models must be able to new variants of behaviors that occur over the course of natural behavior that did not appear in the training data. Will models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time? Assessing the validity of temporally *distal* periods is a crucial step to establish whether automatic classification can be used to measure body position across a day.

In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 34 testing sessions from 22 infants aged 4-14 months. Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each hip) and a video camera to collect ground truth data about infant body position. A *proximal comparison* period began when participants received the equipment and completed a guided phone call during which caregivers were asked to elicit different body positions based on prompts from the experimenter. Although not directly supervised, this period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play while they received instructions from the experimenter. The **first goal** of the current study was to determine the accuracy of body position classification during the proximal comparison period using this novel, semi-supervised procedure in participants' homes. Past work found better performance using "individual models"---models that were trained on one participant's data to predict their later behavior---compared with a "group model" that aggregated data from all infants to create a single body position classifier [@FranchakScott2021], so we compared both modeling approaches in the current investigation.

The crucial test was how well models predicted later behavior over longer recordings of everyday activities. A second, *distal comparison* period followed the proximal comparison period and captured approximately 90 minutes of home behavior that was completely unsupervised. Caregivers and infants could (and did) do whatever they wished. Because this recording happened a considerable amount of time after the initial setup and instructions from the experimenter, accuracy could decline if caregivers or infants moved the garment or sensors. Moreover, increasing variation in everyday activities during the distal comparison creates a greater challenge, testing whether body position classification models can generalize to novel test cases. Thus, the **second goal** of our study was to assess accuracy during the distal comparison.

After the distal comparison period when video recording ceased, we asked caregivers to have infants wear the IMUs for the rest of the day until their regular bedtime, creating the **first real-time, full-day dataset of infant body position**. Interpreting such data required caregivers to log when infants napped, when they removed the sensor garment for diaper changes or other reasons, and when infants went to bed at the end of the day. Thus, the **third goal** of the study was to examine the quality of the full-day data. Could body position be successfully recorded over the desired period? The novel full-day dataset affords us a unique chance to ask whether the estimated time infants spent in different body positions align with past results using video and survey methods. Thus, we determined whether full-day body position measurements conformed to expected age differences in body position. Based on past results [@Survey], infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).

@@

Infants were recruited in one of two age groups: *Younger* infants were between 4 and 7 months and *older* infants were between 11 and 14 months. There were `r num_younger` infants in the 4-7 month group (X female) and `r num_older` in the 11-14 month group (X female). Families were recruited through social media advertisements and from community events in Southern California. Caregivers reported the ethnicity of infants as X Hispanix/Latinx and X for not Hispanix/Latinx. Race was reported as X Caucasian, X Asian, X ~~~ and X others. Families were compensated \$30 for every home recording session they completed. The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study. All caregivers gave their informed consent before the start of the study.
<!-- TODO RACE ETHNICITY SEX, COUNT BY HAND -->

@@

<!-- ## Data Sharing and Transparency -->
<!-- WHAT CAN WE SHARE AND HOW -->
https://nyu.databrary.org/volume/1580 - all raw data files and annotations for one participant

>>>>

authornote: |
  We are grateful to Vanessa Scott, Tasnia Haider, and Ishapreet Kaur for their help in collecting data for the present study and to the research assistants of the UCR Perception, Action, and Development Lab for annotating videos.

  This work was funded by National Science Foundation Grant BCS #1941449 to the first author.

@@

# SUBMITTED DATA IS BASED ON THE FOLLOWING SESSION IDS
#9908  9909  9910  9911 10201 10203 10204 10301 10501 10601 10701 10702 10703 10704 10802 10901 11001 11002 11004 11102 11103 11104 11202 11501 11601 11901 12001 12002 12101 12302 12303 12501 12502 13001
```

@@

In this paper, we present an inertial sensing method to classify infant body position from moment to moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders. We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022], the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a *distal comparison*. Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion [@AiraksinenRasanen2020] and infant-caregiver contact [@YaoPlotz2019]. We provide openly shared data, materials, and analysis code to facilitate other researchers' adopting this method and applying it to populations and behaviors (see Data Sharing and Transparency below).

@@

The LENA&reg; recorder is a commercial device worn in a custom shirt pocket; the recorder has sufficient battery life and storage to record for an entire day. Closed-source LENA&reg; algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the infant. Other long-form audio methods rely on custom-built recorders [@WassPhillips2022], apply alternative classification algorithms to LENA&reg; data [@RasanenSeshadri2020; @MichelettiYao2022], or manually transcribe audio recorded by LENA&reg; devices to improve accuracy or identify behaviors beyond the built-in categories [@Bergelson2019Corpus; @MendozaFausey2021].

Long-form audio recordings have had a transformational impact on language development research by allowing researchers to characterize opportunities for learning in daily life. For example, measuring the amount of speech heard by infants in the home [@Weisleder2013] or in a daycare setting [@PerryPrince2018] revealed how individual differences in speech input predict later vocabulary. Full-day language recording synchronized with other data sources allows researchers to identify how auditory input and vocal production interact with other processes. Beyond individual differences in aggregated data, long-form recordings reveal the temporal schedule of experiences. For example, infants' daily experiences hearing music are clustered in time, with "bursty" episodes of music presence separated by relatively long periods during which music is absent [@MendozaFausey2022]. In addition, synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences. Linking LENA&reg; speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during periods that they were restrained in devices such as swings, exersaucers, and car seats [@MalachowskiSalo2023]. A custom-built wearable ECG and audio recorder allowed @WassPhillips2022 to discover that infant arousal increases the likelihood of infant vocalization across the day.

We identified five key features of long-form audio methods that should be replicated in analogous studies of motor behavior. First, wearable audio recorders are *mobile*. Measurement is not limited to a particular room because the recording device travels with the participant. Data are recorded to onboard device memory, so participants do not need to be in range of a receiver. Second, wearable audio recording is *unobtrusive*. Participants' reactivity to observation, such as from a video camera, may influence behavior. For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA&reg; device [@Bergelson2019Input]. Third and fourth, recordings capture *real-time data* over a *full day*. The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants. Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual [e.g., @WassPhillips2022; @MalachowskiSalo2023]. Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) [@KadookaCaufield2021; @Tamis-LeMondaCustode2018]. "Burstiness" of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent [@BarbaroFausey2022; @WarlaumontSobowale2021]. Fifth, *automatic classification* means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.

But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased. An independent assessment of the LENA&reg; algorithms found mixed results about the accuracy of different measurements. For example, correlations between human transcribed counts of adult words and child vocalizations against LENA&reg;'s automatic counts were strong, *r* = .698 and *r* = .649, respectively [@CristiaLavechin2020]. For other metrics, such as the number of "conversational turns" between the child and communicative partners and the identification of male speakers, agreement was poor. Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day. In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior.

@@

Video and survey methods are the current state-of-the-art in assessing infants' gross motor behavior in naturalistic tasks. Although each method has complementary advantages and disadvantages for characterizing infants' everyday motor experiences, neither method on its own can provide comparable data to the long form audio recordings reviewed in the previous section.

@@

Accordingly, the overarching goal of the current study is to test the validity of long-form body position recording in the home during unsupervised, everyday behavior. Supervised recordings from past work [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022], whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors. Prior work focused on "free play", in which caregivers were asked to play with the infant without restraining the infant or otherwise shaping how they could move. However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position. For example different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to scored as the same behavior. Moreover, classifiers must be able to detect new variants of behaviors that might arise over the course of a real day; it is impossible for researchers to gather training data for every possible variation that might occur. Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time? Assessing the validity of temporally *distal* periods---when infants and caregivers are unsupervised and free to follow their everyday routines---is a crucial step to establish whether automatic classification can be used to measure body position across a day.

In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 34 testing sessions from 22 infants aged 4-14 months. Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each hip) and a video camera to collect ground truth data about infant body position. A *proximal comparison* period began when participants received the equipment and completed a guided phone call during which caregivers were asked to elicit different body positions based on prompts from the experimenter. This "semi-supervised" period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play while they received instructions from the experimenter. The **first goal** of the current study was to determine the accuracy of body position classification during the proximal comparison period using this novel, semi-supervised procedure in participants' homes. Past work found better performance using "individual models"---models that were trained on one participant's data to predict their later behavior---compared with a "group model" that aggregated data from all infants to create a single body position classifier [@FranchakScott2021], so we compared both modeling approaches in the current investigation.

The crucial test was how well models predicted later behavior over longer recordings of everyday activities. A second, *distal comparison* period followed the proximal comparison period and captured approximately 90 minutes of home behavior that was completely unsupervised. Caregivers and infants could (and did) do whatever they wished, and no researcher was present. Because this recording happened a considerable amount of time after the initial setup and instructions from the experimenter, accuracy could decline if caregivers or infants moved the garment or sensors. Moreover, increasing variation in everyday activities during the distal comparison creates a greater challenge, testing whether body position classification models can generalize to novel test cases. Thus, the **second goal** of our study was to assess accuracy during the distal comparison.

After the distal comparison period when video recording ceased, we asked caregivers to have infants wear the IMUs for the rest of the day until their regular bedtime, creating the **first real-time, full-day dataset of infant body position**. Interpreting such data required caregivers to log when infants napped, when they removed the sensor garment for diaper changes or other reasons, and when infants went to bed at the end of the day. Thus, the **third goal** of the study was to examine the quality of the full-day data. Could body position be successfully recorded over the desired period? If full-day classifications of infant behavior are accurate, they should demonstrate convergent validity with other methods. Thus, we determined whether full-day body position measurements conformed to expected age differences in body position that have been identified from previous video and survey methods. Based on past results [@Survey], infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).

@@

Infants were recruited in one of two age groups: *Younger* infants were between 4 and 7 months and *older* infants were between 11 and 14 months. There were `r num_younger` infants in the 4-7 month group and `r num_older` in the 11-14 month group. Ten infants were female and 12 were male. Families were recruited through social media advertisements and from community events in Southern California. Most infants were reported as either Hispanic and White or Non-Hispanic and White. Families were compensated \$30 for every home recording session they completed. The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study. All caregivers gave their informed consent before the start of the study.

@@

## Data Sharing and Transparency
<!-- WHAT CAN WE SHARE AND HOW -->
https://nyu.databrary.org/volume/1580 - all raw data files and annotations for one participant

### 2023-03-27 manuscript.tex
Clean through to methods

Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

Enter author note here.

@@

In this paper, we present an inertial sensing method to classify infant body position from moment to moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders. We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions (Airaksinen et al., 2022, 2020; Franchak et al., 2021), the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings. Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion (Airaksinen et al., 2020) and infant-caregiver contact (Yao et al., 2019).

@@

The LENA® recorder is a commercial device that is worn by infants in a custom shirt pocket; the recorder has sufficient battery life and storage to record for an entire day. Closed-source LENA® algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the infant. Other long-form audio methods rely on custom-built recorders (Wass et al., 2022), apply alternative classification algorithms to LENA® data (Micheletti et al., 2022; Räsänen et al., 2020), or manually transcribe audio recorded by LENA® devices to improve accuracy or identify behaviors beyond the built-in categories (Bergelson, Casillas, et al., 2019; Mendoza \& Fausey, 2021).

Long-form audio recordings have had a transformational impact on language development research by allowing researchers to characterize opportunities for learning in daily life. For example, measuring the amount of speech heard by infants in the home (Weisleder \& Fernald, 2013) or in a daycare setting (Perry et al., 2018) revealed how individual differences in speech input predict later vocabulary. Full-day language recording synchronized with other data sources allows researchers to identify how auditory input and vocal production interact with other processes. Beyond individual differences in aggregated data, long-form recordings can be used to determine the temporal schedule of experiences. For example, infants' daily experiences hearing music are clustered in time, with ``bursty'' episodes of music presence separated by relatively long periods during which music is absent (Mendoza \& Fausey, 2022). In addition, synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences. Linking LENA® speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during periods that they were restrained in devices such as swings, exersaucers, and car seats (Malachowski et al., 2023). A custom-built wearable ECG and audio recorder allowed Wass et al. (2022) to discover that infant arousal increases the likelihood of infant vocalization across the day.

We identified five key features of long-form audio methods that should be replicated in analogous studies of motor behavior. First, wearable audio recorders are \emph{mobile}. Measurement is not limited to a particular room because the recording device travels with the participant. Data are recorded to onboard device memory, so participants do not need to be in range of a receiver. Second, wearable audio recording is \emph{unobtrusive}. Participants' reactivity to observation, such as from a video camera, may influence behavior. For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA® device (Bergelson, Amatuni, et al., 2019). Third and fourth, recordings capture \emph{real-time data} over a \emph{full day}. The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants. Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual (e.g., Wass et al., 2022). Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) (Kadooka et al., 2021, April; Tamis-LeMonda et al., 2018). ``Burstiness'' of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent (Barbaro \& Fausey, 2022; Warlaumont et al., 2021). Fifth, \emph{automatic classification} means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription. But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased. An independent assessment of the LENA® algorithms found mixed results about the accuracy of different outcomes. For example, correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively (Cristia et al., 2020). For other metrics, such as the number of ``conversational turns'' between the child and communicative partners and the identification of male speakers, agreement was poor.

Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day. In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior.

@@

Infants were recruited in one of two age groups: \emph{Younger} infants were between 4 and 7 months and \emph{older} infants were between 11 and 14 months. There were 8 infants in the 4-7 month group (X female) and 14 in the 11-14 month group (X female). Families were recruited through social media advertisements and from community events in Southern California. Caregivers reported the ethnicity of infants as X Hispanix/Latinx and X for not Hispanix/Latinx. Race was reported as X Caucasian, X Asian, X \textasciitilde\textasciitilde\textasciitilde{} and X others. Families were compensated \$30 for every home recording session they completed. The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study. All caregivers gave their informed consent before the start of the study.

>>>>

We are grateful to Vanessa Scott, Tasnia Haider, and Ishapreet Kaur for their help in collecting data for the present study and to the research assistants of the UCR Perception, Action, and Development Lab for annotating videos.

This work was funded by National Science Foundation Grant BCS \#1941449 to the first author.

@@

In this paper, we present an inertial sensing method to classify infant body position from moment to moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders. We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions (Airaksinen et al., 2022, 2020; Franchak et al., 2021), the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a \emph{distal comparison}. Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion (Airaksinen et al., 2020) and infant-caregiver contact (Yao et al., 2019). We provide openly shared data, materials, and analysis code to facilitate other researchers' adopting this method and applying it to populations and behaviors (see Data Sharing and Transparency below).

@@

The LENA® recorder is a commercial device worn in a custom shirt pocket; the recorder has sufficient battery life and storage to record for an entire day. Closed-source LENA® algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the infant. Other long-form audio methods rely on custom-built recorders (Wass et al., 2022), apply alternative classification algorithms to LENA® data (Micheletti et al., 2022; Räsänen et al., 2020), or manually transcribe audio recorded by LENA® devices to improve accuracy or identify behaviors beyond the built-in categories (Bergelson, Casillas, et al., 2019; Mendoza \& Fausey, 2021).

Long-form audio recordings have had a transformational impact on language development research by allowing researchers to characterize opportunities for learning in daily life. For example, measuring the amount of speech heard by infants in the home (Weisleder \& Fernald, 2013) or in a daycare setting (Perry et al., 2018) revealed how individual differences in speech input predict later vocabulary. Full-day language recording synchronized with other data sources allows researchers to identify how auditory input and vocal production interact with other processes. Beyond individual differences in aggregated data, long-form recordings reveal the temporal schedule of experiences. For example, infants' daily experiences hearing music are clustered in time, with ``bursty'' episodes of music presence separated by relatively long periods during which music is absent (Mendoza \& Fausey, 2022). In addition, synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences. Linking LENA® speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during periods that they were restrained in devices such as swings, exersaucers, and car seats (Malachowski et al., 2023). A custom-built wearable ECG and audio recorder allowed Wass et al. (2022) to discover that infant arousal increases the likelihood of infant vocalization across the day.

We identified five key features of long-form audio methods that should be replicated in analogous studies of motor behavior. First, wearable audio recorders are \emph{mobile}. Measurement is not limited to a particular room because the recording device travels with the participant. Data are recorded to onboard device memory, so participants do not need to be in range of a receiver. Second, wearable audio recording is \emph{unobtrusive}. Participants' reactivity to observation, such as from a video camera, may influence behavior. For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA® device (Bergelson, Amatuni, et al., 2019). Third and fourth, recordings capture \emph{real-time data} over a \emph{full day}. The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants. Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual (Malachowski et al., 2023; e.g., Wass et al., 2022). Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) (Kadooka et al., 2021, April; Tamis-LeMonda et al., 2018). ``Burstiness'' of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent (Barbaro \& Fausey, 2022; Warlaumont et al., 2021). Fifth, \emph{automatic classification} means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.

But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased. An independent assessment of the LENA® algorithms found mixed results about the accuracy of different outcomes. For example, correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively (Cristia et al., 2020). For other metrics, such as the number of ``conversational turns'' between the child and communicative partners and the identification of male speakers, agreement was poor. Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day. In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior.

@@

Infants were recruited in one of two age groups: \emph{Younger} infants were between 4 and 7 months and \emph{older} infants were between 11 and 14 months. There were 8 infants in the 4-7 month group and 14 in the 11-14 month group. Ten infants were female and 12 were male. Families were recruited through social media advertisements and from community events in Southern California. Most infants were reported as either Hispanic and White or Non-Hispanic and White. Families were compensated \$30 for every home recording session they completed. The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study. All caregivers gave their informed consent before the start of the study.
